\chapter*{Preface}
\setheader{Preface}

Long ago, in 2013, I performed my B.Sc. thesis under Dr. Jos Thijssen. He chose the most educational course, which was to let me make errors. This is, by far, the most effective way to be taught, even if it is not the most effective way to success.

When I asked Dr. Thijssen if he had M.Sc. thesis subjects in 2014, he had a whole list. We settled on this particular subject, which is to take the non-equilibrium Green's Function formalism and add capacitive interactions to it analytically.  This was exciting; it was theoretical work promising publication at some point. Additionally, we would try to explain the measurements made by Dr. Mickael Perrin, which were significantly weaker than the theory apparently predicted.

The thesis, I think, succeeded, despite the fact that Dr. Thijssen spent most of his attention on educating other minds, whereas I also had a Ph.D. grant proposal to write. Even so, we looked at a large portion of the parameter fields of our relatively simple model, trying to understand its intricacies. I also found a promising derivation for incorporating phononic interactions.

Of course, Dr. Thijssen does not work alone; conversations with his Ph.D. student, Jose Celis Gil, were very helpful and intriguing. Likewise, I've also spoken to Dr. Mickael Perrin and Max Koole several times, which I think greatly helped understanding the experimental side of the story. 

So, I would like to thank my advisor, Dr. Jos Thijssen, and his Ph.D. student, Jose Celis Gil. Without their experience, knowledge, guidance and participation this thesis would not have been possible. Additionally, I want to thank Dr. Jos Seldenthuis, whom I have never met but whose work and notes are the very foundation this thesis builds upon.

For mental support and listening to my ramblings, I'd like to thank my family, Maaike Jans (girlfriend), Elske Hottinga (mother), Dick de Boer (father), Roy de Boer (brother), Rick de Boer (brother). Also several of my friends, Elizabeth Berghuijs, Jacob Verhaart, Bram van der Veen, Marije Barel and Esmeralda Tomas\"oa.


\begin{flushright}
{\makeatletter\itshape
    \@firstname\ \@lastname \\
    Delft, \today
\makeatother}
\end{flushright}

